---
title: "Final Project Presentation"
author: "Group 15: Annie Cao, Turner Goetsch, Jiuru Lyu, Ben Zamler, Zhenyu Zou"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(ggmap)
library(lubridate)
library(maps)
library(plotly)
library(shiny)
library(tidyverse)
```

```{r, eval=FALSE, include=FALSE}
data <- read_delim("data.txt", delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  drop_na() %>%
  filter(user_location_country == "UNITED STATES OF AMERICA" & site_name == "EXPEDIA.COM" & is_mobile == 0 & is_package == 0 & hotel_country == "UNITED STATES OF AMERICA") %>%
  filter(prop_is_branded == 1) %>%
  filter(prop_starrating == 4 | prop_starrating == 5) %>%
  mutate(srch_ci = ymd(srch_ci, tz = "America/New_York"), 
         srch_co = ymd(srch_co, tz = "America/New_York")) %>%
  filter(srch_ci %within% interval(ymd_hms("2015-06-01 00:00:00 EDT"), ymd_hms("2015-08-31 23:59:59 EDT")) & srch_co %within% interval(ymd_hms("2015-06-01 00:00:00 EDT"), ymd_hms("2015-08-31 23:59:59 EDT")))
```

```{r, eval=FALSE, include=FALSE}
write.csv(data, "new_data.csv")
```

```{r, include=FALSE}
# When re-opening this file, only run chunks after this. 
data <- read.csv("new_data.csv")
dest <- read_delim("dest.txt", delim = "\t", escape_double = FALSE, trim_ws = TRUE)
state_pops <- read.csv("us-state-populations.csv")
party <- read.csv("party.csv")
```

Data Inclusion Criteria:

-   US users
-   using `Expendia.com`
-   not from mobile site
-   not from a package research
-   searching for a domestic hotel
-   hotel is branded
-   hotel has a star rating above 4
-   check-in date from June 1st, 2015 to Augest 31st, 2015
-   check-out date from June 1st, 2015 to Augest 31st, 2015

## Shiny App on the distribution of `popular_food_`

```{r, echo=FALSE}
food <- read.csv("new_data.csv") %>% 
  filter(is_booking == 1) %>%
  inner_join(dest, by = "srch_destination_id") %>%
  pivot_longer(col = starts_with("popular_food"), names_to = "popular", values_to = "probability") %>%
  select(popular, probability) %>%
  mutate(probability = 10 ^ probability * 100)

names <- food %>%
  distinct(popular)
```

```{r, echo=FALSE}
ui <- fluidPage(
    titlePanel("Compare Three Popular Scores"),
    selectInput(inputId = "popular1", 
                label = "Popular Score 1", 
                choices = names), 
    selectInput(inputId = "popular2", 
                label = "Popular Score 2", 
                choices = names), 
    selectInput(inputId = "popular3", 
                label = "Popular Score 3", 
                choices = names), 
    mainPanel(plotlyOutput(outputId = "boxplot"))
)

server <- function(input, output) {
  output$boxplot <- renderPlotly({
    food %>%
      filter(popular == input$popular1 | popular == input$popular2 | popular == input$popular3) %>%
        plot_ly(x = ~popular, y = ~probability, type = "box", color = ~popular, boxpoints = FALSE)
    })
}

shinyApp(ui = ui, server = server)
```

## Map showing the per mil of number of bookings per population in different states of US.

```{r, echo=FALSE}
map_data <- map_data("state") %>%
  mutate(region = toupper(substr(region, 1, 2))) %>%
  distinct(region, .keep_all = TRUE)
```

```{r, echo=FALSE}
map <- data %>%
  group_by(user_location_region) %>%
  mutate(n = sum(is_booking)) %>%
  select(user_location_region, n) %>%
  distinct(user_location_region, .keep_all = TRUE) %>%
  left_join(state_pops, by=c( "user_location_region" = "code")) %>%
  mutate(rate = n / pop_2014 * 1000) %>%
  left_join(party, by = c("user_location_region" = "state_code")) %>%
  select(user_location_region, rate, party)
```

```{r, echo=FALSE}
plot_geo(map, locationmode = "USA-states") %>%
  add_trace(locations = ~user_location_region, type = "choropleth", z = ~rate, text = ~party, colorscale = "Reds")
```

## A heat map comparing `distance_band` and `hist_price_band` with regard to the number of bookings.

```{r, echo=FALSE}
dat <- data %>%
  select(distance_band, hist_price_band, is_booking) %>%
  group_by(distance_band, hist_price_band) %>%
  count(is_booking) %>%
  filter(is_booking == 1)

dat$distance_band <- factor(dat$distance_band, levels = c("VC", "C", "M", "F", "VF"))
dat$hist_price_band <- factor(dat$hist_price_band, levels = c("VL", "L", "M", "H", "VH"))

dat %>%
  plot_ly(x = ~factor(distance_band), y = ~factor(hist_price_band), z = ~n, colorscale = "Reds") %>%
  add_heatmap()
```
