---
title: "HW01_Jiuru Lyu"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1: Import the dataset "fruitfly" 
```{r}
fruitfly <- read.csv('fruitfly.csv')
head(fruitfly)
```

# Question 2: Provide a simple plot to investigate the distribution of "lifespan" in the dataset.
```{r}
hist(fruitfly$lifespan, main='Histogram of Fruitfly Lifespan', xlab='Lifespan')
```

```{r}
plot(fruitfly$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan', xlab = 'Index', ylab = 'Lifespan')
```

However, the data set is grouped by type, so I also would like to generate separate plots for each type. 

To do so, I firstly create sub data sets according to the types of the fruit fly: 
```{r}
fruitfly_type1 <- fruitfly[26:50,]
fruitfly_type2 <- fruitfly[51:75,]
fruitfly_type3 <- fruitfly[1:25,]
fruitfly_type4 <- fruitfly[76:100,]
fruitfly_type5 <- fruitfly[101:125,]
```

## Fruitfly Type1
```{r}
hist(fruitfly_type1$lifespan, main='Histogram of Fruitfly Lifespan - Type1', xlab='Lifespan')
```
```{r}
plot(fruitfly_type1$No, fruitfly_type1$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan - Type1', xlab = 'No', ylab = 'Lifespan')
```

## Fruitfly Type2
```{r}
hist(fruitfly_type2$lifespan, main='Histogram of Fruitfly Lifespan - Type2', xlab='Lifespan')
```
```{r}
plot(fruitfly_type2$No, fruitfly_type2$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan - Type2', xlab = 'No', ylab = 'Lifespan')
```

## Fruitfly Type3
```{r}
hist(fruitfly_type3$lifespan, main='Histogram of Fruitfly Lifespan - Type3', xlab='Lifespan')
```

```{r}
plot(fruitfly_type3$No, fruitfly_type3$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan - Type3', xlab = 'No', ylab = 'Lifespan')
```

## Fruitfly Type 4
```{r}
hist(fruitfly_type4$lifespan, main='Histogram of Fruitfly Lifespan - Type4', xlab='Lifespan')
```


```{r}
plot(fruitfly_type4$No, fruitfly_type4$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan - Type4', xlab = 'No', ylab = 'Lifespan')
```

## Fruitfly Type 5
```{r}
hist(fruitfly_type5$lifespan, main='Histogram of Fruitfly Lifespan - Type5', xlab='Lifespan')
```

```{r}
plot(fruitfly_type5$No, fruitfly_type5$lifespan, type = 'p', main = 'A Plot of Fruitfly Lifespan - Type5', xlab = 'No', ylab = 'Lifespan')
```

# Question 3:Generate a HTML document of your rmd and submit on Canvas.