---
title: "HW04"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Packages
```{r}
library(tidyverse)
library(babynames)
```

## Question 1

**For **`babynames` **data set, find the top 5 most popular names that are shared by both males and females.**
```{r, message = FALSE, warning = FALSE}
babynames %>%
  group_by(name) %>%
  summarize(total = sum(n)) %>%
  top_n(5)
```

The top 5 most popular names are **James**, **John**, **Marry**, **Michael**, and **Robert**. 

## Question 2

**Pick two different years at your choice to see their top 5 most popular shared names. Are they the same?**

 I would like to investigate the top 5 most popular shared names in the years of 1893, 1902。

```{r, message = FALSE, warning = FALSE}
babynames %>%
  group_by(year, name) %>%
  filter(year == 1893 | year == 1902) %>%
  summarize(total = sum(n)) %>%
  top_n(5)
```

As indicated by the result: 

1. The top 5 popular names in 1893 are: Anna, James, John, Mary, and William. 

2. The top 5 popular names in 1902 are: Helen, James, John, Mary, and William. 

3. The names *James*, *John*, *Mary*, and *William* are shared by the two years. 

