---
title: "HW02_Jiuru Lyu"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before starting solving the homework questions, I am importing the `gss` data set. The `gss` data set has been stored in the sample work directory as this .Rmd file.

```{r}
gss <- read.csv('gss.csv')
```

# Question 1

***Create a small random sample of size 1000 from gss data set. (set.seed(1234)), name it "smalldat" and save/export on your current working directory.***

To find exactly how many rows of data am I selecting from, I firstly find the dimension of the `gss` data set.

```{r}
dim(gss)
```

The result shows that there are 53474 rows in total, so I will get 1000 random numbers from 1 to 53474 without replacement and set them to the row numbers of extraction.

```{r}
set.seed(1234)

smalldat <- gss[sample(1:53474, 1000, replace = FALSE), ]

head(smalldat)
```

Now, I am saving the data into my working directory.

```{r}
write.csv(smalldat, 'smalldat.csv')
```

The `smalldat` data set is successfully saved.

# Question 2

***Calculate the average age of `smalldat`.***

To calculate the average, we can use the `mean()` function.

```{r}
mean(smalldat$age)
```

The result is weird: R returns `NA`, indicating either there is a missing value or some data is not numeric. I suspect that there is at least a missing value in our random sample, so I run `is.na()` function to testify if any value is missing in the age column of `smalldat` data set.

```{r}
is.na(smalldat$age)
```

As indicated by the result, there are some missing values in our random sample. That explains why we cannot get a numerical value for the average mean if we do not specify the value of `na.rm` argument in the `mean()` function. To successfully finding the average age and stripping the `NA` values, I run the following code:

```{r}
mean(smalldat$age, na.rm = TRUE)
```

Now the average age is successfully calculated, ignoring the existence of the missing values. The result tells us that the average mean of our random sample is `r mean(smalldat$age, na.rm = TRUE)`.

# Question 3

***Provide an appropriate plot to investigate the distribution of degree of `smalldat`.***

Previously, we have run the `view(smalldat)` code, and I found that the degree column in the data set is a categorical variable specifying the degree of education the responders have received. Hence, bar plot would be the most appropriate way to show the distribution of degree.

Before creating the bar plot, I need to create the frequency table for the degree column first.

```{r}
degree_frequency <- table(smalldat$degree)

degree_frequency
```

Using the `degree_frequency` table, I can now create the bar plot:

```{r}
barplot(degree_frequency, main = 'Bar Plot of the Degree Column in Smalldat', xlab = 'Types of Degree', ylab = 'Number of Responders')
```

# Question 4 

***Calculate the percentage of married participants with bachelor degree in smalldat (among all married participants, the percentage of them with bachelor degree)?***

To answer this question, a proportion contingency table is created using the `prop.table()` and `table()` functions. Because the question asks me to find "the percentage of them with bachelor degree," a column marginal proportion table is created by setting the `margin` argument equal to 1. 
```{r}
prop.table(table(smalldat$degree, smalldat$marital), margin = 2)
```

As the table indicates, the result of the first row and the the second column indicates the answer we are seeking, which is `r prop.table(table(smalldat$degree, smalldat$marital), margin = 2)[1, 2]` (i.e., there are 14.2% of married participants holding a bachelor degree in our random sample).

# Question 5

***Provide an appropriate plot to investigate the relationship between marital status and education status (degree) in smalldat.***

As the martial and degree columns in our data set are all categorical variables, I need to create a bar plot to represent their relationship. In this case, I will not use the marginal proportion table anymore; rather, I will create a frequency table of two variables.
```{r}
barplot(table(smalldat$degree, smalldat$marital), main = 'Relationship Between Marital and Education Status in Smalldat',  xlab = 'Marital Status', ylab = 'Number of Responders', legend.text = TRUE)
legend = c('BACHELOR','GRADUATE',' HIGH SCHOOL', 'JUNIOR COLLEGE', 'LT HIGH SCHOOL' )
```

