---
title: "HW03"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import `gss` data set:
```{r}
gss <- read.csv('gss.csv')
```

# Question 1

- **How many missing values for the variable marital?**

```{r}
sum(is.na(gss$marital))
```

Hence, there are 20 missing values fro the variable marital. 

- **What is the percentage of female respondents (excluding missing values), try function prop.table()?**

```{r}
prop.table(table(gss$sex))
```

The percentage of female respondents is 0.5614317 (56.1%).

- **What is the percentage of male respondents who felt “NOT TOO HAPPY”?**

```{r}
prop.table(table(gss$sex, gss$happy), margin = 1)
```

The percentage of male respondents who felt "NOT TOO HAPPY" is 0.1203201 (12.0%).

- **What is the percentage of married respondents who age between 30 and 40 (including the boundaries)?**

As `age` column is a numeric column, I need to convert it into character values before creating the prop.table. 
```{r}
gss$age1 <- factor(NA, levels = c("Below 30", "Between 30 and 40", "Above 40"))

gss$age1[gss$age < 30] <- "Below 30"
gss$age1[gss$age >= 30 | gss$age <=40] <- "Between 30 and 40"
gss$age1[gss$age > 40] <- "Above 40"
```

Let's check if the factor is created correctly: 
```{r}
summary(gss$age1)
```

Now, we are comfortable to create our prop.table. 
```{r}
prop.table(table(gss$marital, gss$age1), margin = 1)
```
Hence, the percentage of married respondents who age between 30 to 40 (including boundaries) is 0.41880282 (41.9%). 

- **What is the average age of female married respondents who took the survey from 1999 and 2010?**

Firstly, we want to select the relevant data that we need. 
```{r}
marriedF <- gss$age[gss$sex == "FEMALE" & gss$marital == "MARRIED"]

marriedF_99_10 <- marriedF[gss$year >= 1999 | gss$year <= 2010]
```

Now, we can calculate the average age. Because there are missing values, we set `na.rm = TRUE`.
```{r}
mean(marriedF_99_10, na.rm = TRUE)
```

Therefore, the average age of female married respondents who took the survey from 1999 and 2010 is 44.59293. 

# Question 2

- **Make a summary of year. What type of vector is the year variable?**

```{r}
summary(gss$year)
```

The year variable is a numeric variable. 

- **It should be treated as a factor variable. Change it into a new factor variable YEAR in the gss data. Make a new summary of YEAR, which year had the most participants?**

```{r}
gss$year1 <- factor(NA, levels = as.character(c(1973:2010)))

for (i in 1973:2010) {
  gss$year1[gss$year == i] <- as.character(i)
}

summary(gss$year1)
```

To find out which year had the most participants, I would like to create a bar plot. 
```{r}
barplot(table(gss$year1))
```

Apparently, year 2006 had the largest number of participants, and the numbe of participants in 2006 was 4510. 

- **Use the function tapply to get average age of respondents by YEAR. What years had the oldest and the youngest average age? (look up `tapply()`)**

Firstly, I would like to loop up how to use the function `tapply()`. 
```{r}
?tapply()
```

Now, I am confident to use `tapply()`. I will now start solving this question. 
```{r}
tapply(gss$age, gss$year1, mean, na.rm = TRUE)
```

Hence, the year 1984 had the youngest average age, which is 44.00477, and the year 2010 had the oldest average age, which is 48.04207. 

# Question 3

- **Reorder the levels in degree of gss data based on the number of years in education. **

Firstly, we need to conver the degree column into a factor. 
```{r}
gss$degree_factor <- as.factor(gss$degree)
```

Let's check the levels and the order of the levels: 
```{r}
levels(gss$degree_factor)
```

We want the levels in an order of the years of education, so we reorder it in the following way: 
```{r}
gss$degree_factor <- factor(gss$degree, levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))
```

Now, let's double-check the order: 
```{r}
levels(gss$degree_factor)
```

- **Provide the appropriate plot to visualize the distribution of the new degree variable.**

The best way to visualize factor and character data is bar plot: 
```{r}
barplot(table(gss$degree_factor))
```


