---
title: "Introduction to R"
author: "Jiuru Lyu"
date: "2022-08-25"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r}
# Loading packages
library('tidyverse')
```

# The use of R and RStudio

## The Four Panels in RStudio

You can directly type your codes in the script. Using cmd+return to run a line of the code.

```{r}
2+3
```

You can also plot something using R.

```{r}
hist(treering)
```

Practice: Installing packages - knitr and rmarkdown

If you have any questions regarding a function in R, you can use the `help` function to look for help, or you can use `?function_name`.

```{r}
help('hist')
```

```{r}
?mean
```

Also, Google is also a very good tool when you need help.

## The Very Basic R Programming Syntax

1.  R can do simple calculations

-   Addition, Subtraction, Multiplication, and Division

```{r}
2+3
```

```{r}
2-3
```

```{r}
2*3
```

```{r}
2/5
```

-   Finding the reminder of one number divided by another:

```{r}
6%%4
```

-   Finding the quotient (the whole number part) of a division

```{r}
6 %/% 4
```

2.  You can use the **assign operator** to assign a value to a variable.

```{r}
# Assigning the value 1 to x
x <- 1
x
```

3.  There will be many objects in R and RStudio. Some objects are preset, such as the data set `trees`.

```{r}
trees
```

If we want to access a specific column of a preset data set, we can use the `$` operator.

```{r}
trees$Girth
```

Another typical object in R is a vector. We use the `c()` function to create a vector. The syntax is `c(starting number:ending number)`

```{r}
v <- c(1:5)
v
```

4.  Functions in R There are many pre-set functions in R. To call them, just type them in the code script.

```{r}
mean(trees$Height)
```

```{r}
hist(trees$Height)
```

When you are unsure about how to use a specific function, we can use the `help()` function.

```{r}
help(hist)
```

```{r}
?hist
```

If you want to look up for a function called `bs()`, we will not get any results if you type `?bs`, this is because the function is included in an unloaded package. To look up the function in all packages, we use the following code:

```{r}
help("bs", try.all.packages = TRUE)
```

## R Markdown Introdunction

A R Markdown follows all syntax of markdown.

1.  *italics* and **bold**

2.  `verbatim code`

3.  ~~strikethrough~~

4.  $A=\pi *r^{2}$

5.  Use `#` to indicate titles.

6.  Use `<>` to add a link. Use `[R Markdown](http://rmarkdown.rstudio.com)` to add a hyperlink (The result will be [R Markdown](http://rmarkdown.rstudio.com) after the file is knitted).

7.  Use `![Name](File.name)` to include an image.

8.  To make a list use `number` or `-`.

9.  To create a table: 

| Right | Middle | Left |
|:------|:-------|:-----|
| 1     | 3      | 1    |
| 2     | 2      | 2    |

When setting up a code chunk, we can set what the ouput will include.

```{r, echo=FALSE, evak=FALSE, message=FALSE}
# this is an example code chunk
```

-   The `echo = FALSE` parameter is to prevent printing of this R code chunk.
-   The `eval=FALSE` parameter is to control if the code chunk will print the R result.
-   The `include=FALSE` parameter is to prevent printing of both R code and R result.
-   The `message=FALSE` parameter is to prevent printing of the warning or errors in R console.

You can also run some simple inline R codes using the syntax `r print('your code here')`. You will note see the results of this inline code until you knit the markdown document into a HTML file. All code starts with `r` to tell the R Markdown that you are including a R code.

## Importing a Data Set
1. Method 1: Import data set from Environment panel -> copy the code here from console panel
```{r， echo = FALSE}
gss <- read.csv("~/Library/CloudStorage/OneDrive-EmoryUniversity/2 Freshman/1 Fall 2022/QTM 150/Labs/Lab_1/gss.csv")
```
2. Method 2: use working directory to import date set: files -> data set -> import data -> copy the working directory -> read.XXX
3. Method 3: make sure download/save dataset and current rmd document are in the same folder -> get the working directory
```{r}
getwd()
gss<-read.csv("gss.csv")
```
