---
title: "Data Manipulation - dplyr"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Data Set and Packages
```{r}
gss <- read.csv("newgss.csv")
```
```{r}
library(tidyverse)
library(babynames)
```


# `dplyr` Introduction

`dplyr` provides simple *verbs*, functions that correspond to the most common data manipulation tasks, to help you translate your thoughts into code.

`dplyr` aims to provide a function for each basic verb of data manipulation:

- filter() to select cases based on their values.
- arrange() to give cases/observations a specific order
- select() to select variables based on their names.
- mutate() to add new variables that are functions of existing variables.
- group_by() to split the data frame into pieces based on some variables.
- summarize() to condense multiple values to a single value.

*split-apply-combine*: A common analytical pattern is to split data frame into pieces, apply some function to each pieces, and combine the results back together again to a new data frame. 

All verbs work similarly:
1. The first argument is a data frame;
2. The subsequent arguments describe what to do with the data frame, using the variable names (without quotes);
3. The result is a new data frame

## The Pipe Operator: `%>%`
Pipes take the output from one function and feed it to the first argument of the next function.
Instead of nesting functions (reading from the inside to the outside), the idea of of piping is to read the functions from left to right.

## The `filter()` Verb

We use `filter()` to select cases/observations/rows based on their values.
```{r}
# Select all the respondents in 2000
gss %>%
  filter(year == 2000) %>%
  head()
```
```{r}
# Select all the respondents in 2000 and 2001
gss %>%
  filter(year == 2000 | year == 2001) %>%
  head()
```
```{r}
# Select all the respondents in 2000 and 2001 who are married
gss %>%
  filter(year == 2000 | year == 2001, marital == "MARRIED") %>%
  head()
```

We can store our data using the assign operator: 
```{r}
married_00_01 <- gss %>%
  filter(year == 2000 | year == 2001, marital == "MARRIED")
```

## The `arrange()` Verb
`arrange()` reorders the observations by their values of a variable.

```{r}
gss %>%
  arrange(age) %>%
  head()
```

The default order for `arrange()` is ascending order. To see descending order, we use `desc()`.
```{r}
gss %>%
  arrange(desc(age)) %>%
  head()
```

We can also arrange the data frame using two or more variables. Usually, the first variable entered will be given the higher priority when arranging.
```{r}
gss %>%
  arrange(desc(age), year) %>%
  head()
```


`filter()` and `arrange()` work on **observations/rows**

## The `select()` Verb
We use `select()` to select variables based on their names.
```{r}
gss %>%
  select(year, age, marital) %>%
  head()
```
```{r}
gss %>%
  select(3:7) %>%
  head()
```
```{r}
gss %>%
  select(-c(finrela)) %>%
  head()
```

## The `mutate()` Verb

`mutate()` creates new variables/columns or alters pre-existing variables/columns
```{r}
gss %>%
  mutate(
    var1 = income1 / age
  ) %>%
  head()
```

If we only want to show the new variable we create, we can use `transmute()`
```{r}
gss %>%
  transmute(var1 = income1 / age) %>%
  head()
```

`select()` and `mutate()` work on **variables/columns** 

## The `summarize()` Verb
We use `summarize()` to get a summary of variables/columns
```{r}
gss %>%
  summarize(
    AvgAge = mean(age, na.rm = TRUE),
    AvgIncome = mean(income1, na.rm = TRUE)
  )
```

## Use Pipe
```{r}
gss %>%
  select(year, income1, age, sex) %>%
  filter(year == 2000, sex == "MALE") %>%
  arrange(age) %>%
  mutate(var1= income1 / age) %>%
  head()
```
```{r}
gss %>%
  filter(year == 2000, marital == "MARRIED") %>%
  select(3:6) %>%
  arrange(desc(age)) %>%
  summarize(Avg = mean(age, na.rm = T))
```

## The `group_by()` Verb

`group_by()` splits the entire dataframe according to the groups. 
```{r}
gss %>%
  group_by(sex) %>%
  summarize(Avg = mean(age, na.rm = T))
```

*split-apply-combine*: split data frame into groups, apply some function to each group, and combine the results back together again to a new data frame. 

***Example***: Create a ggplot to visualize the trend of average income1 for different sex over years.
```{r}
gss %>%
  group_by(sex, year) %>%
  summarize(
    AvgIncome = mean(income1, na.rm = T)
  ) %>%
  ggplot(aes(x = year, y = AvgIncome, color = sex)) + 
    geom_line() + 
    xlab("Year") + 
    ylab("Average Income") + 
    ggtitle("Average Income Over Years, Grouped by Sex") + 
    theme_bw()
```

***Example***: Get the average of all respondents each year and visualize it
```{r}
gss %>%
  group_by(year) %>%
  summarize(AveAge = mean(age, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = AveAge)) + 
    geom_line() + 
    theme_bw()
```

# `dplyr` Review

## `gss` Data Set

***Exercise 1***: For the `gss` data, get the percent of married respondents by year, Get a plot to visualize the marital status by year.

$\boxed{\text{Method 1}}$
```{r}
gss %>%
  group_by(year) %>%
  summarize(prec = sum(marital == "MARRIED", na.rm = TRUE) / sum(!is.na(marital))) %>%
  head()
```

$\boxed{\text{Method 2}}$
```{r}
gss %>%
  group_by(year) %>%
  summarize(prec = prop.table(table(marital))[2]) %>%
  ggplot(aes(x = year, y = prec)) + 
    geom_line()
```

***Exercise 2***: For the `gss` data, get a data about the percent of married respondents and average age of respondents by year, then provide a plot to visualize the relation between age and marital status by year.
```{r}
gss %>%
  group_by(year) %>%
 summarize(prec = prop.table(table(marital))[2], 
           AvgAge = mean(age, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = AvgAge, size = prec, color = prec)) + 
    geom_point()

```

***Exercise 3***: For the `gss` data, find the percent of respondents with \$25,000 or more by party affiliation and year.
```{r}
gss %>%
  group_by(partyid, year) %>%
  summarize(prec = sum(income == "$25000 OR MORE", na.rm = TRUE) / sum(!is.na(income)) * 100) %>%
  head()
```

***Exercise 4***: What is the "richest" region? - does it change over the years?

$\boxed{\text{Method 1}}$
```{r}
gss %>%
  group_by(year, region) %>%
  summarize(AvgIncome = mean(income1, na.rm = TRUE)) %>%
  filter(AvgIncome == max(AvgIncome)) %>%
  ggplot(aes(x = year, y = AvgIncome, color = region, size = region)) +
    geom_point()
```

$\boxed{\text{Method 2}}$
```{r}
gss %>%
  group_by(year, region) %>%
  summarize(AvgIncome = mean(income1, na.rm = TRUE)) %>%
  group_by(year) %>%
  summarize(riches = max(AvgIncome), 
            region = region[which.max(AvgIncome)]) %>%
  ggplot(aes(x = year, y = riches, color = region, size = region)) +
    geom_point()
```

***Extension***: Which region has the lowest income in the year of 1980? 
```{r}
gss %>%
  group_by(year, region) %>%
  summarize(AvgIncome = mean(income1, na.rm = TRUE)) %>%
  group_by(year) %>%
  summarize(riches = min(AvgIncome), 
            region = region[which.min(AvgIncome)]) %>%
  filter(year == 1980)
```

## `babynames` Data Set

```{r}
str(babynames)
```
```{r}
summary(babynames)
```

***Exercise 5*** Create a new variable to see the rank of names from the most popular to the least popular.
```{r}
babynames %>%
  mutate(rank = min_rank(n)) %>%
  arrange(desc(rank)) %>%
  head()
```
Note: `min_rank()` ranks the variable from the minimum to the maximum (i.e., the one with the lowest value is ranked number 1. )

***Extension*** Find the most popular name in 2017. 
```{r}
babynames %>%
  mutate(rank = min_rank(n)) %>%
  arrange(desc(rank)) %>%
  filter(year == 2017)
```

***Exercise 6*** For `babynames` data set, visualize the trend of born children population over years.
```{r}
babynames %>%
  group_by(year) %>%
  summarize(total = sum(n)) %>%
  ggplot(aes(x = year, y = total)) +
    geom_line()
```

***Exercise 7*** Find the number of unique names in the year 2017. 
```{r}
babynames %>%
  group_by(name) %>%
  filter(year == 2017) %>%
  summarize(total = sum(n)) %>%
  count(name) %>%
  filter(n == 1) %>%
  summarize(total = sum(n))
```

Note: `count(variable) = group_by(variable) %>% summarize(n())` (the number of observations in each group). `n()` is the number observations in each group. `count()` is different from `sum(numeric variable)`. 

***Exercise 8*** Provide an appropriate to visualize the top 5 popular names for both girl and boy.

$\boxed{\text{Method 1}}$
```{r}
babynames %>%
  group_by(sex, name) %>%
  summarize(total = sum(n)) %>%
  top_n(5) %>%
  ggplot(aes(x = name, y = total)) +
    geom_bar(stat = "identity", aes(fill = sex))
```
Note: `geom_col() = geom_bar(stat = "identity")`

$\boxed{\text{Method 2}}$
```{r}
babynames %>%
  group_by(sex, name) %>%
  summarize(total = sum(n)) %>%
  top_n(5) %>%
  ggplot(aes(x = name, y = total)) +
    geom_bar(stat = "identity") + 
    facet_grid(.~sex, scale = "free")
```

***Exercise 9*** The top 5 most popular names in 2000, see their trend over years
Finding the 5 most popular names in 2000: 
```{r}
namesIn2000 <- babynames %>%
  filter(year == 2000) %>%
  group_by(name) %>%
  summarize(total = sum(n)) %>%
  top_n(5) %>%
  select(name)
```

Converting `namesIn2000` to vector. 
```{r}
as.vector(namesIn2000)

class(as.vector(namesIn2000))

class(as.vector(namesIn2000)[[1]])
```

Complete the exercise
```{r}
babynames %>%
  filter(name %in% namesIn2000[[1]]) %>%
  group_by(year, name) %>%
  summarize(total = sum(n)) %>%
  ggplot(aes(x = year, y = total, color = name)) + 
    geom_line()
```

# Join

## Pakcage: nycflights13

We will use the `nycflights13` package to learn about relational data. `nycflights13` contains four data tables that are related to the main `flights` table:

```{r}
library(nycflights13)
```

The `flights` data set is the main data set. 
```{r}
head(flights)
```

Relational data sets: 

- `plane`connects to the main data set via a single variable `tailnum`
```{r}
head(planes)
```

- `airports` connects to the main data setvia two variables `dest` = `faa`
```{r}
head(airports)
```

- `airlines` connects to the main data set through the `carrier` variable 
```{r}
head(airlines)
```

- `weater` connects to the main data set via `origin`, `year`, `month`, `day`, and `hour`
```{r}
head(weather)
```

## Definitions

The variables used to connect each pair of tables are called *keys*. A key is a variable (or set of variables) that uniquely identifies an observation.

A *primary key* variable uniquely identify each observation

***Example*** syaing each `tailnum` only has one observation (`tailnum` is the primary key variable in planes) is equivalent to say that `tailnum` is the foreign key in flights (uniquely identifies an observation in another table)
```{r}
planes %>%
  group_by(tailnum) %>%
  summarize(amount = n()) %>%
  filter(amount != 1)
```

- `n()` is to calculate the number of occurrences of each observation

- `count()` is to calculate the number of occurrences of each observation, the same as the below ```group_by()+summarize(n())```
```{r}
planes %>%
  count(tailnum) %>%
  filter(n != 1)
```

## Joins
In `dplyr` there are two types of join functions, mutating joins and filtering joins.

To learn the two types of joins, we will use the following data set: 
```{r}
band <- tribble(
   ~name,     ~band,
  "Mick",  "Stones",
  "John", "Beatles",
  "Paul", "Beatles"
)


instrument <- tribble(
    ~name,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)

instrument1 <- tribble(
    ~artist,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)
```


### Mutating Joins

```XXX_join(x, y, by = "key variable")```

-   join y to x, x is the primary data set

-   mutate new variables in x by copying different variables in y

-   return a new data set

-   `XXX_join(x, y, by = c("key variable1", "key variable2"))`: A key is a variable that uniquely identifies an observation, otherwise, we need multiple variables to identify an observation

1. *Left join*: `left_join(x, y, by = "key variable")` keeps all the observations (rows) in x
```{r}
left_join(band, instrument, by = "name")
```

2. *Right join*: `right_join(x, y, by = "key")` keep all the observations (rows) in y
```{r}
right_join(band, instrument, by = "name")
```

Use can use `setequal` to test if two data sets are identitical.

3. *Inner join*: `inner_join()` keeps all the observations in **both** x and y: 
```{r}
inner_join(band, instrument, by = "name")
```

Be careful when using `inner_join` because we will lose information in this case. 

4. *Full join*: `full_join()` keeps **all** observations in x and y: 
```{r}
full_join(band, instrument, by = "name")
```

### Filtering Joins

1. *Semi join*: `semi_join(x,y, by="key")` keeps all the observations of x that have a match in y: 
```{r}
semi_join(band, instrument, by = "name")
```

***Exercise***: Use `right_join`, `left_join`, `filter()`, `select()` to get the same dataset above with `semi_join()`

Note: we can check the code line by line to see what to do next. 

$\boxed{\text{Method 1}}$
```{r}
band %>%
  left_join(instrument, by = "name") %>%
  filter(!is.na(plays)) %>%
  select(name, band)
```

$\boxed{\text{Method 2}}$
```{r}
band %>%
  right_join(instrument, by = "name") %>%
  filter(!is.na(band)) %>%
  select(name, band)
```

```{r}
flights %>%
  count(dest) %>%
  top_n(10)
```

2. *Anti Join*: `anti_join(x,y, by="key")` drops all the observations of x that have a match in y. `anti_join()` also provides a great way to diagnose joins that go wrong.

***Example***: Use an anti_join() to return the rows of artists for which you don't have any instrument info.
```{r}
anti_join(band, instrument, by = "name")
```

## Set Operations

When two data sets contain the same variables and each variable is used in the key: we can use the folloing set operations: 

- `union(band, instrument)`:  these two data sets have different variables

In the following set operations requires `data_set1` and `data_set2` have the same variables

-` intersect(data_set1, data_set2)`

- `setdiff(data_set1, data_set2)`: observations in the `data_set1` but not the `data_set2`

- `setequal()`:  check two data sets with the same observations in any order

- `identical()`:  check two data sets with the same observations in the same order

- `cbind()`: Add more variables

- `rbind()`: Add more rows to a data set. 

The following two operations are exactly the same as the two above, but they are faster to use. 

- `bind_cols()`  

- `bind_rows()`

***Example***
```{r}
band1 <- tribble(
  ~name, ~surname,
  "John",   "Lennon",
  "Paul",   "McCartney",
  "George", "Harrison", 
  "Ringo",  "Starr"
)

band2 <- tribble(
   ~name,     ~surname,
  "Mick",    "Jagger",
  "Keith",   "Richards",
  "Charlie", "Watts",
  "Ronnie",  "Wood"
)
```

Combine band1 and band2: 
```{r}
bind_rows(band1, band2)
```

```{r}
rbind(band1, band2)
```

- Add a new column to indicate which observations (rows) are from: 
```bind_rows(label names for observations in new column, .id=column names for new column)```

```{r}
bind_rows(Beatles=band1, Stones=band2, .id="band")
```

## Examples in the `nycflight13` Package
***Example 1***: How many flights without airline information in the airlines? 
```{r}
anti_join(flights, airlines, name = "carrier")
```

Because we got 0 rows from the code above, all the flights in the `flights` data set have information in the `airline` data set. 

***Extension***: Swtich the order: 
```{r}
airlines %>%
  anti_join(flights, name = "carrier")
```

We still get 0 rows, which means all information in the airlines has information in the flights. Hence, the flights and airlines data sets are exactly matched. 

How many different airlines in the flights and airlines?

$\boxed{\text{Method 1}}$
```{r}
flights %>%
  count(carrier)
```

We get 16 rows, so there are 16 different ariline carriers. 

$\boxed{\text{Method 2}}$
```{r}
flights %>%
  summarize(n_distinct(carrier))
```

***Example 2***: How many flights without airport information in the airport? 
```{r}
flights %>%
  anti_join(airports, by = c("dest" = "faa")) %>%
  summarize(n_distinct(dest))
```

**Note: in the flights data set, the variable `dest` corresponds to the `faa` variable in the airports data set. So, we need to use `by = c('var1' = 'var2')` in the joins. **

What are these airports?
```{r}
flights %>%
  anti_join(airports, by = c("dest" = "faa")) %>%
  count(dest)
```

Hence, the four airports are BQN, PSE, SJU, and STT. 

***Example 3***: Find the top 10 popular airports in flights. 
```{r}
flights %>%
  count(dest) %>%
  top_n(10)
```

***Extension***: If we want the results to be sorted: 
```{r}
flights %>%
  count(dest) %>%
  arrange(n) %>%
  top_n(10)
```

```{r}
flights %>%
  count(dest, sort = T) %>%
  head(10)
```
**Note: Here, we can directly use `head(10)` to select the first 10 rows because the data set has already been sorted. **

How many different airlines have flight to those top 10 destinations? 
```{r}
top10_dest <- flights %>%
  count(dest) %>%
  top_n(10)
```
```{r}
flights %>%
  semi_join(top10_dest, by = "dest") %>%
  summarize(n_distinct(carrier))
```

Hence, there are 13 carriers having flights to those top 10 destinations. 

***Example 4***: Find the top 5 popular names shared by girls and boys each year. 
```{r}
babynames %>%
  count(year, name) %>%
  filter(n == 2) %>% # Find the names that are shared by both girls and boys each year. 
  select(year, name) %>%
  left_join(babynames, by = "name") %>% # Join with the original data set, and then find the top 5 names. 
  group_by(year.x, name) %>%
  summarize(total = sum(n)) %>%
  top_n(5)
```

Connecting flights and planes to find how many distinct airplanes in flights that don’t have information in planes?

```{r}
flights %>%
  anti_join(planes, by = "tailnum") %>%
  summarize(n_distinct(tailnum))
```


