---
title: "5 Data cleaning - tidyr"
author: "Jiuru Lyu"
date: '`r Sys.Date()`'
output: 
  html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to `tidyr`
```{r}
library(tidyverse)
```


The `tidyr` package helps tidy up messy data sets.

There are three interrelated rules which make a data set tidy:

-   Each variable must have its own column

-   Each observation must have its own row

-   Each value must have its own cell

There are four key functions in the `tidyr` package, `pivot_longer()`, `pivot_wider()`, `separate()`, `unite()`.


# Pivoting

There are two common problems in a untidy dataset:

-   One variable might be spread across multiple columns.

-   One observation might be scattered across multiple rows.

To fix these problems, you will need the two most important functions in tidyr: `pivot_longer()` and `pivot_wider()`.

## `pivot_longer()`

A common problem is a data set where some of the column names are not names of variables, but *values* of a variable.

`tibble` is very similar to `data.frame`, which is better to read a large data set.
```{r}
table1 <- tibble(
  `country` = c("Afghanistan","Brazil","China"),
  `1999`    = c(19987071, 172006362, 1272915272),
  `2000`    = c(20595360, 174504898, 1280428583)
)

table1
```

```{r}
table2 <- tibble(
  `country` = c("Afghanistan","Brazil","China"),
  `1999`    = c(745, 37737, 212258),
  `2000`    = c(2666, 80488, 213766)
)

table2
```

To tidy a data set like `table1` and `table2`, we need to pivot the offending columns into a new pair of variables, it makes "wide" data longer.

Template:

```
pivot_longer(dataset,
              a set of columns = "The set of columns whose names are values, not variables", 
              names_to = "The name of the variable to move the column names to", 
              values_to = "The name of the variable to move the column values to"
            )
```

```{r}
table1 %>%
  pivot_longer(c("1999", "2000"), names_to = "year", values_to = "population")
```

```{r}
table2 %>%
  pivot_longer(c("1999", "2000"), names_to = "year", values_to = "case")
```

`pivot_longer()` makes datasets longer by increasing the number of rows and decreasing the number of columns relatively.

## `pivor_wider()`

`pivot_wider()` is the opposite of `pivot_longer()`. We use it when an observation is scattered across multiple rows. `pivot_longer()` makes wide tables narrower and longer, and `pivot_wider()` makes long tables shorter and wider.

```{r}
table3<- tibble(
  `country` = c("Afghanistan", "Afghanistan", "Afghanistan", "Afghanistan", "Brazil", "Brazil", "Brazil", "Brazil","China", "China","China", "China"),
  `year`    = c(1999,1999,2000,2000,1999,1999,2000,2000,1999,1999,2000,2000),
  `type`    = c("case", "population","case", "population","case", "population","case", "population","case", "population","case", "population"),
  `count`   = c(745, 19987071, 2666, 20595360, 37737, 172006362, 80488, 174504898, 212258, 1272915272, 213766, 1280428583)
)

table3
```

Template: 

```
pivot_wider(dataset, 
              names_from = "the column to take variable names from", 
              values_from = "the column to take values from"
            )
```

```{r}
table3 %>%
  pivot_wider(names_from = "type", values_from = "count")
```

# Separate and Unit

`separate()` pulls apart one column into multiple columns, `unite()` is the inverse of `separate()`.

```{r}
table4<-tibble(
  `country` = c("Afghanistan", "Afghanistan", "Brazil", "Brazil","China", "China"),
  `year`    = c(1999,2000,1999,2000,1999,2000),
  `rate`    = c("745/19987071", "2666/20595360", "37737/172006362", "80488/174504898", "212258/1272915272", "213766/1280428583")
)

table4
```

## `separate()`

Template: 
```
separate(dataset, old variable, into = c("",""), sep = "")
```

You can separate by a notation, such as `/`, `,`, etc. 
```{r}
table4 %>%
  separate(rate, into = c("case", "population"), sep = "/")
```

You can also separate according to the position (count from the left): 
```{r}
table4 %>%
  separate(rate, into = c("case", "population"), sep = 2)
```

## `separate()` and `unite()`
```{r}
library(nycflights13)
```

```{r}
flights %>%
  separate(time_hour, into = c("date1", "hour1"), sep=" ") %>% # from left-most
  head(10)
```

```
unite(dataset, new variable, a sets of existing columns)
```
```{r}
flights %>%
  unite(date, year, month, day) %>%
  head(10)
```

```{r}
flights %>%
  unite(date, year,month, day) %>%
  separate(time_hour, into=c("date1","hour1"), sep=10) %>% # from left-most
  head(10)
```

# Missing Values
```{r}
stocks <- tibble(
  year   = rep(c(2015,2016), each=4),
  qtr    = rep(c(1,2,3,4), 2),
  return = c(1.88, 0.59, 0.35, NA, NA, 0.92,0.17, 2.66)
)

stocks
```

We can use `na.omit()` to make missing values implicit
```{r}
stocks %>%
  na.omit()
```

We can use `complete()` to make missing values explicit
```{r}
stocks %>%
  na.omit() %>%
  complete(year, qtr)
```

We can also use `fill` to fill the missing values with the most recent non missing vlaue. 
```{r}
stocks %>%
  fill(return)
```


# Examples
***Example 1***: 
```{r}
stocks1 <- tibble(
  `half` = c(1,2),
  `2015` = c(1.88, 0.59),
  `2016` = c(0.92, 0.17)
)

stocks1
```

```{r}
stocks1 %>%
  pivot_longer(c("2015", "2016"), names_to = "year", values_to = "return")
```

***Example 2***: 
```{r}
cases <- tribble(
  ~Country, ~"2011", ~"2012", ~"2013",
      "FR",    7000,    6900,    7000,
      "DE",    5800,    6000,    6200,
      "US",   15000,   14000,   13000
)

cases
```

```{r}
cases %>%
  pivot_longer(c("2011":"2013"), names_to = "year", values_to = "cases")
```


***Example 3***: 
```{r}
pollution <- tribble(
       ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",     121,
   "Beijing", "small",     121
)

pollution
```

```{r}
pollution %>%
  pivot_wider(names_from = "size", values_from = "amount")
```



