---
title: "6 Dealing with Categorical Variables - forcats"
author: "Jiuru Lyu"
date: "`r Sys.Date()`"
output: 
 html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Packages and Data Sets
```{r}
library(tidyverse)
```

```{r}
glimpse(gss_cat)
```



## Introudciton to `forcasts`

This package provides some tools for working with categorical variables (factors)

***Example***: find the relation between religions and average hours of watching tv.

```{r}
relig_summary <- gss_cat %>%
  na.omit(tvhours) %>%
  group_by(relig) %>%
  summarize(avg = mean(tvhours))
  
ggplot(relig_summary, aes(x = relig, y = avg)) + 
  geom_bar(stat = "identity") + 
  coord_flip()
```

How could we interpret the plot without overall pattern? i.e., can we re-arrange this plot so that it is in descending or scending order? 

## `fct_reorder()`

Reordering the levels of a factor using `fct_reorder()`.

`fct_reorder()` takes three arguments:

- f, the factor whose levels you want to modify.

- x, a numeric vector that you want to use to reorder the levels.

- Optionally, fun, a function used if there are multiple values of x for each value of f. The default value is median.

```
fct_reorder(cat, numeric)
```

```{r}
relig_summary %>%
  ggplot(aes(x = avg, y = fct_reorder(relig, avg))) +
    geom_col()
```

We can use fact_rev to reverse the order of the factor. 
```
fct_rev(fct_reorder(cat, numeric))
```

We can also use pip: 
```
fct_reorder(cat, numeric) %>% fct_rev()
```

```{r}
relig_summary %>%
  ggplot(aes(x = avg, y = fct_reorder(relig, avg) %>% fct_rev())) + 
    geom_col()
```

## `fct_relevel`

Now, we want to alwasy keep `Don't know` at the bottom of our plot, no matter we are using ascending or descending order. 

You can use `fct_relevel()`. It takes a factor, f, and then any number of levels that you want to move to the front of the line (the bottom of the plot).

```f
ct_relevel(cat, "level", after=)
```

```{r}
relig_summary %>%
  ggplot(aes(x = avg, y = relig %>% fct_reorder(avg) %>% fct_relevel("Don't know", after = 0))) + 
    geom_col()
```

## `fct_infreq()`

You can use `fct_infreq()` to order levels in increasing frequency: this is the simplest type of reordering because it does not need any extra variables. You may want to combine with `fct_rev()`.


Note in the following graph, marital is not ordered. 
```{r}
gss_cat %>%
  ggplot(aes(x = marital)) + 
    geom_bar()
```

Using `fact_infreq()`, we can make it ordered
```{r}
gss_cat %>%
  ggplot(aes(x = fct_infreq(marital))) + 
    geom_bar()
```

We can use `sort()` to creat a basic R plot that is ordered. 
```{r}
barplot(sort(table(gss_cat$marital)))
```

To reverse the order, we can combined `fct_infreq()` with `fct_rev()`.
```{r}
gss_cat %>%
  ggplot(aes(x = marital %>% fct_infreq() %>% fct_rev())) + 
    geom_bar()
```

## `fct_recode`

***Example***: Find the relation between party affiliation and average hours of watching tv.

```{r}
gss_cat %>%
  group_by(partyid) %>%
  summarize(avg = mean(tvhours, na.rm = T))
```

But it is still hard to answer the question because the `partyid` is not clearly defined. We need to recode factor `fct_recode()`

Let's check the levels of `partyid` first: 
```{r}
levels(factor(gss_cat$partyid))
```

Now, we can change the levels using `fct_recode()`: 
```
fct_recode(cat, )
```

```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  mutate(partyidnew = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party")) %>%
  group_by(partyidnew) %>%
  summarize(avg = mean(tvhours, na.rm = T))
  
```

## `fct_collapse()`

If you want to collapse a lot of levels, `fct_collapse()` is a useful variant of `fct_recode()`. For each new variable, you can provide a vector of old levels:

```{r}
gss_cat %>%
  mutate(partyidnew = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  ))  %>%
  group_by(partyidnew) %>%
  summarize(avg = mean(tvhours, na.rm = T))
```

## `fct_lump()`

Sometimes you just want to lump together all the small groups to make a plot or table simpler. That is the job of `fct_lump()`.

```{r}
gss_cat %>%
  mutate(relignew = fct_lump(relig)) %>%
  count(relignew)
```

but we have probably over collapsed:

```{r}
gss_cat %>%
  mutate(relignew = fct_lump(relig,n=5)) %>%
  count(relignew)
```

## `case_when`

```{r}
mtcars %>%
  mutate (
    transmission =
      case_when(
        am==0 ~ "automatic",
        am==1 ~ "manaul"
      )
  ) %>%
  head(10)
```

```{r}
mtcars %>%
  mutate (
    gear_char =
      case_when(
        gear==3 ~ "three",
        gear==4 ~ "four",
        gear==5 ~ "five"
      )
  ) %>%
  head(10)
```

```{r}
starwars %>%
  select(name:mass, gender, species) %>%
  mutate(
    type = case_when(
      height > 200 | mass > 200 ~ "large",
      species == "Droid"        ~ "robot",
      TRUE                      ~ "other"
    )
  ) %>%
  head(10)
```

***Example***: For gss_cat, create a plot to look at how average age varies across income level (rincome) with the order.
```{r}
gss_cat %>%
  drop_na(age) %>%
  mutate(rincome = fct_recode(rincome,
    "other" = "No answer", 
    "other" = "Don't know", 
    "other" = "Refused", 
    "other" = "Not applicable",
  )) %>%
  group_by(rincome) %>%
  summarize(avg = mean(age)) %>%
  ggplot(aes(x = avg, y = rincome %>% fct_reorder(avg) %>% fct_relevel("other", after = 0))) +
    geom_col() + 
    labs(
      x = "Income Levels",
      y = "Average Age",
      title = "Average Age for Different Income Levels for gss_cat Data Set"
    )
```

