---
title: "Flow Control and Function"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Flow Control
Some sontrol statements affect or control the execution of other statements in some way.

## `ifelse()` Function'
Syntax: 
```
ifelse(testexpression, yes, no)
```

***Example 1***: Convert a character vector to a numeric vector
```{r}
x <- c("female", "male", "male", "female", "female", "male", "female")

ifelse(x == "female", 0, 1)
```

***Example 2***: Convert uniform random numbers to binomial random numbers.

`rXXX(size, parameters of distribution)` create random numbers from some specific probability distribution
```{r}
set.seed(23)
x <- runif(10,0,6) # Create random numbers from a uniform distribution from (0,6)

ifelse(x > 3, 1, 0)

```

***Example 3***: Import `gss_cat` dataset, create a new variable `racenew` only with values `0` and `1`, where `1` represents white, and `0` represents non-white.
```{r}
gss_cat %>%
  mutate(racenew = ifelse(race == "White", 1, 0)) %>%
  head(10)
```

## `if() else` statement
The `if() else` statement allows us to control which statements are executed. 

Syntax: 
```
if (condition) {commands when condition is TRUE}
else {commands when condition is FALSE}
```

```{r}
x <- 5

if (x < 0) {
  print("negative")
} else {
  print("non-negative")
}
```

Sometimes, when there is only one command to execute, we can omi `{}`.

We can also omit `print()`.
```{r}
x <- c("female", "male")

if (is.numeric(x)) "numeric vector" else "non-numeric vector"
```

AND operator
```{r}
x <- 15

if (x > 0 & x < 10) {
  print("target")
} else {
  print("error")
}
```

OR operator
```{r}
x <- -5

if (x > 0 | x < 0) {
  print("non-zero number")
} else {
  print("zero")
}
```

NOT operator
```{r}
if (!FALSE){
  cat("this statement will show up")
}
```

## Nested `if() else` statement
Syntax: 
```
if (test_expression1) {
  statement 1
} else if (test_expression2) {
  statement 2
} else if (test_expression3) {
  statement 3
} else {
  statement 4
}
```

```{r}
x <- 15

if (x<0) {
  print("negative number")
} else if (x > 0) {
  print("positive number")
} else {
  print("zero")
}
```

***Example***: Print the letter grade for a grade of one student in QTM 150. If grade > 90, we give them a A; if 80 < grade < 90, we give them a B; otherwise C.
```{r}
set.seed(111)
grade <- rnorm(1, 50, 50)

if (grade > 90) {
  print("A")
} else if (grade > 80 & grade <= 90) {
  print("B")
} else if (grade > 70 & grade <= 80) {
  print("C")
} else {
  print("D")
}
```

# The `for` Loop
The `for` loop controls how many times the statement are repeated. 

If we want to run the same block of code multiple times, the `for()` loop statement allows one to specify that a certain operation should be repeated a fixed number of times. 

## Iteration `n` times over an index
***Example 1***: Calculate the sum from 1 to 100.
```{r}
n <- 100
x <- 0 # Initial value

for (i in 1:n) {
  x <- x + i
}
x
```

***Example 2***: Adding stuff to a vector
```{r}
set.seed(1234)
random_numbers <- NULL # Create a vector with no elements

for (i in 1:10) {
  random_numbers <- c(random_numbers, runif(1)) # runif(1) creates random numbers from a uniform distribution in [0,1]
}

cat(random_numbers)

```

## Iterate over all items in a vector
***Example 1***: In R, we have some US state facts and figures. Find all the states starting with the letter `M`.
```{r}
Mstates <- NULL

for (x in state.name) {
  if (substr(x, 1, 1) == "M") {
    Mstates <- c(Mstates, x)
  }
}
Mstates
```

***Example 2***: Central Limit Theorem
**Central Limit Theorem (CLT)**: the distribution of sample means approximates a normal distribution as the sample size gets larger, regardless of the population's distribution. i.e., The sampling distribution (histogram) of sample means (selection with random sample size) from a (fixed) population is approximately normal. 
```{r}
pop <- 1:100
size <- 10
iteration <- 10000
sample <- NULL
```

$\boxed{\text{Method 1}}$: Use the `for()` Loop
```{r}
set.seed(1108)

for (i in 1:iteration) {
  x <- mean(sample(pop, size, replace = FALSE))
  sample <- c(sample, x)
}

hist(sample, breaks = 50)
```

$\boxed{\text{Method 2}}$: We can also use `replicate()` function. The syntax goes follows: 
```
replicate(#of\repetition, expression)
```
```{r}
set.seed(1108)

hist(replicate(iteration, mean(sample(pop, size))), breaks = 50)
```

***Example 3***: get the average grades for each student
```{r}
# Look at the data set
set.seed(1000)

exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)

exams
```

If we run `mean(exams)`, we will get an error. 
```{r}
mean(exams)
```

This is because, we need `[[]]` to extract the values in the vector format instead of single bracket: 
```{r}
exams[[1]]
```

If we use single bracket `[]`, we will get the result not in the vector format. The result will be in the list format. 
```{r}
exams[1]
```


What we can do: 
```{r}
list(
  student1 = mean(exams$student1), 
  student2 = mean(exams$student2), 
  student3 = mean(exams$student3), 
  student4 = mean(exams$student4), 
  student5 = mean(exams$student5)
)
```

However, when the number of students gets larger, this method is very not efficient. 

$\boxed{\text{Method 1}}$: We can use `for()` loop instead.
```{r}
mean_score = NULL

for (i in seq_along(exams)){
  mean_score = c(mean_score, mean(exams[[i]]))
}

mean_score
```

`seq_along()` gives a vector including all the index of the argument. 

$\boxed{\text{Method 2}}$: We can also use `Xapply()` function. 
```{r}
lapply(exams, mean) # lapply will return another list
```
```{r}
sapply(exams, mean) # sapply will return the result in vecotr format 
```

***Extension***: Find the range of grades that each student get from the exams. 
```{r}
range = NULL

for (i in seq_along(exams)) {
  range <- c(range, max(exams[[i]]) - min(exams[[i]]))
}

range
```


# Function

We have been using many functions in R, now we learn to write them ourselves. We will use the `function()` function to create functions.
Syntax: 

- Functions that include one input and one output

```
function_name <- function(input) {
  Body of the function
  Output value (return)
}
```

- Functions that include multiple inputs and outputs

```
function_name <- function(input1, input2, input3, ...) {
  Body of the function
  return(list = c(output1, output2, output3, ...))
}
```

- When recalling the function, we simple use
```
function_name(input)
```

***Example 1***: Create a mean function
```{r}
my_mean <- function(x) {
  m <- sum(x) / length(x)
  return(m)
}

my_mean(c(1:10))
```

Functions can be declared and used inside a function. Functions in R can also return functions.

***Example 2***: Create a mean function for inputs including `NA` values
```{r}
my_mean <- function(x) {
  if (sum(is.na(x)) > 0) {
    x <- na.omit(x)
  }
  m <- sum(x) / length(x)
    return(m)
}

my_mean(c(1:5, NA))

```

Or, we can let the user to choose if they want the `NA` to be omitted or not. 
```{r}
my_mean <- function(x, na.rm) {
  if (na.rm == TRUE) {
    if (sum(is.na(x)) > 0) {
      x <- na.omit(x)
    }
  }
  m <- sum(x) / length(x)
  return(m)
}

my_mean(c(1:5, NA), na.rm = TRUE)
my_mean(c(1:5, NA), na.rm = FALSE)
```

***Example 3***: Get some summaries of a numeric vector
```{r}
summaries <- function(x) {
  return(list = c(min(x), max(x), mean(x), median(x), sd(x), length(x)))
}

summaries(c(1:15))
```

***Example 4***: Make a function to calculate the range
```{r}
my_range <- function(x) {
  r <- max(x) - min(x)
  return(r)
}

my_range(1:100)
```

Or we can combine the `range()` function in R to write `my_range()` function
```{r}
my_range <- function(x) {
  r <- range(x)[2] - range(x)[1]
  return(r)
}

my_range(1:100)
```

***Example 5***: Make a re-scale function that re-scale all the numbers in a vector to [0,1]
We can use the following function to re-scale: 
$$x_r=\frac{x-\min{x}}{\max{x}-\min{x}},$$
where $x_r$ is the re-scaled $x$ value. 
```{r}
rescale <- function(x) {
  (x-min(x)) / my_range(x)
}

rescale(1:10)
```

***Example 6***: Plot a function
$\boxed{\text{Method 1}}$
```{r}
func1 <- function(x) { 
  x^2
}

curve(func1,-1,1)
```

$\boxed{\text{Method 2}}$
```{r}
fun2<-function (x) {
  y<-x^2+2*x+1
  plot(x,y)
}

fun2(seq(-1,1,0.01))
```

$\boxed{\text{Method 3}}$
```{r}
ggplot(data=data.frame(x=c(-2,2)), aes(x=x))+
  stat_function(fun=func1)+ 
  xlim(c(-2,2))
```

We can also use `optimize()` to find the minimum for the function at (0,0)
```{r}
optimize(func1, interval=c(-2,2))
```

***Example 7***
```{r}
myfun <- function(x) {
  1 / (1 + exp(-x + 10))
}
curve(myfun(x), from = 0, to = 20)
curve(1 - myfun(x), add = TRUE, col = "red") # Add a line
```


***Example 8***: Create a function to print all the summaries
Note: If we do not want to use `return(list = c(output1, output2, output3,...)`, we can also use `print()` to print each output. 
```{r}
summaries <- function (vector) {
  print(length(vector))
  print(summary(vector))
  print(mean(vector,na.rm = T))
  print(sd(vector,na.rm=T))
  print(quantile(vector))
  par(mfrow=c(1,2))
  hist(vector)
  boxplot(vector)
}

x <- rnorm(100)
summaries(x)
```


# The `purrr` Package

The function in `purrr` makes iteration easier.

Recall the data set including students exam results over the semester. 
```{r}
# rXXX(size, parameters of this distribution) to create random numbers from a specific probability distribution
set.seed(1000)
exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)
exams

extra_credit <- list(0, 0, 10, 10, 15)
```

We can make a function to calculate many students' exams average
```{r}
collection_mean <- function(list) {
  output <- vector("double", length(list))  # double vector, is a numeric vector
  for (i in seq_along(list)) { # seq_along() is to get index of elements of the list, 1:length(list) works as well
    output[i] <- mean(list[[i]])
  }
  output
}

collection_mean(exams)
```

## `map()` function in the purrr package

The goal of using `purrr` functions instead of for loops is to allow you to break common list manipulation challenges into independent pieces:

How can you solve the problem for a single element of the list? Once you’ve solved that problem, purrr takes care of generalizing your solution to every element in the list. With `purrr`, you get lots of small pieces that you can compose together with the pipe.

`map()` function for each type of output:

- `map()` makes a list.
- `map_lgl()` makes a logical vector.
- `map_int()` makes an integer vector.
- `map_dbl()` makes a double vector.
- `map_chr()` makes a character vector.

Each function takes a vector/list as input, applies a function to each piece, and then returns a new vector/list that’s the same length

***Example 1***
```{r}
# map(vector/list, function)
map(exams, mean) # map the mean function to each element, and then return a list

map_dbl(exams, mean) # return a vector
```

***Example 2***: Calculate the variance (`var()`) of each student exam grades.
```{r}
map(exams, mean)

map(exams, sd)

exams %>% 
  map(var)

exams %>%
  map(max)
```

***Example 3***: Calculate the max grade (`max()`)for each student. Return the result as a vector.
```{r}
exams %>% 
  map_dbl(max)  # double vector, numeric vector
```

***Example 4***: Calculate the range for each student. Return the result as a vector.
```{r}
range <- function(x) {
  max(x) - min(x)
}
exams %>%
  map_dbl(range)
```

***Example 5***:  Compute a final grade for each student, where the final grade is the average test score plus any `extra_credit` assigned to the student. Return the results as a double (i.e. numeric) vector.
$\boxed{\text{Method 1}}$
```{r}
final_grade <- function(list, y) {
  mean(list) + y
}

map2_dbl(exams, extra_credit, final_grade)
```

***Extension***: Compute a final grade for each student, where the final grade is the highest test score plus any `extra_credit` assigned to the student
```{r}
final_grade <- function(list, y) {
  max(list) + y
}

map2(exams, extra_credit, final_grade)
```

$\boxed{\text{Method 2}}$
```{r}
exams %>%
  map(mean) %>%
  map2(extra_credit, sum)
```

$\boxed{\text{Method 3}}$ The most efficient way. 
```{r}
map2_dbl(exams, extra_credit, function(x ,y) {mean(x) + y})
```

## `map()` works on a dataset

***Example 1***: Compute the mean of every column in `mtcars`

$\boxed{\text{Method 1}}$
```{r}
map_dbl(mtcars,mean)
```

$\boxed{\text{Method 2}}$
```{r}
.x<-mtcars[[1]]      # map understands . specially
mean(.x)            # solve for single .x
```

```{r}
map_dbl(mtcars, ~ mean(.x)) # generalize solution with map() function
```

***Example 2***: Fit the linear model for each piece of data.
```{r}
mtcars %>%
  split(.$cyl) %>%
  map(function(df) lm(mpg~wt, data = df))
```

***Shortcut***: `.` refers to the current list element (in the same way that i referred to the current index in the for loop).

```{r}
mtcars %>%
  split(.$cyl) %>%
  map(~lm(mpg~wt, data = .))
```

***Example 3***:
```{r}
mtcars %>%
  split(.$cyl) %>%
  map(~ggplot(., aes(wt, mpg))+geom_point())
```

***Example 4***: Compute the number of unique values in each column.
```{r}
.x<-iris[[1]]
length(unique(.x))
```

```{r}
map(iris, ~ length(unique(.x)))
```

***Example 5***: Fix the missing values: 
```{r}
df<- tibble(
  a= c(rnorm(9), -99),
  b<-c(-999, -99, rnorm(8))
)

df
```

```{r}
fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}
```

```{r}
map(df, fix_missing)
```



